Block bootstrap consistency\\under weak assumptions

### 2015-12-29 Paper.tex
Write initial version of block bootstrap theory paper

\documentclass[11pt]{article}
\usepackage{amsfonts}
\usepackage{amsmath,amsthm,amssymb,graphicx,setspace,url,booktabs,tabularx,slantsc}
\usepackage[T1]{fontenc}

% \usepackage[yyyymmdd,hhmmss]{datetime}
% \usepackage{fancyhdr}
% \pagestyle{fancy}
% \lfoot{Page \thepage. Typeset on \today\ at \currenttime}
% \cfoot{}
% \rfoot{}

\usepackage[sort,round]{natbib}
\usepackage[multiple]{footmisc}
\usepackage{geometry}
\usepackage[small]{caption}

@@

\newtheorem{thm}{Theorem}
\newtheorem{lem}{Lemma}
\newtheorem{claim}{Claim}
\newtheorem{cor}{Corollary}
\newtheorem{res}{Result}

@@

\DeclareMathOperator{\dist}{d}
\DeclareMathOperator{\const}{const}
\DeclareMathOperator{\cov}{cov}
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\eqd}{\overset{d}{=}}
\DeclareMathOperator{\ind}{1}
\DeclareMathOperator{\pr}{Pr}
\DeclareMathOperator{\sgn}{sgn}
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\vech}{vech}

@@

\newcommand{\clt}{\textsc{clt}}
\newcommand{\fclt}{\textsc{fclt}}
\newcommand{\hac}{\textsc{hac}}
\newcommand{\lln}{\textsc{lln}}
\newcommand{\mds}{\textsc{mds}}
\newcommand{\ned}{\textsc{ned}}

\frenchspacing

@@

\author{Gray Calhoun\thanks{Economics Department, Iowa State
    University, Ames, IA 50011. Telephone: (515) 294-6271.  Email:
    \texttt{gcalhoun@iastate.edu}. Web: \texttt{http://gray.clhn.co}.
    I would like to thank Helle Bunzel, Dimitris Politis, Robert
    Taylor, and three anonymous referees for their comments and
    feedback on earlier versions of this paper.}}

\title{Block bootstrap consistency\\under weak assumptions}

\date{March 25, 2013}
\maketitle

\begin{abstract}\noindent
  This paper weakens the size and moment conditions needed for typical
  block bootstrap methods (i.e. the Moving Blocks, Circular Blocks,
  and Stationary Bootstraps) to be valid for the sample mean of
  Near-Epoch-Dependent (\ned) functions of mixing processes; they are
  consistent under the weakest conditions that ensure the original
  \ned\ process obeys a Central Limit Theorem \citep[those
    of][\textit{Econometric Theory}]{Jon:97}.  In doing so, this paper
  extends de Jong's method of proof, a blocking argument, to hold with
  random and unequal block lengths.  This paper also proves that
  bootstrapped partial sums satisfy a Functional CLT under the same
  conditions.

@@

\newpage
\noindent Block bootstraps, e.g. the Moving Blocks
\citep{Kun:89,LiS:92}, Circular Block \citep{PoR:92}, and Stationary
Bootstraps \citep{PoR:94}, have become popular in Economics, partly
because they do not require the researcher to make parametric
assumptions about the data generating process.  They are valid under
general weak dependence and moment conditions.  Some recent papers
(\citealp{GoW:02}; \citealp{GoJ:03}) relax the dependence and moment
conditions of the original papers to fit with those commonly used in
Econometrics based on Near-Epoch-Dependence
(\ned).\footnote{\citet{GoW:02} show that these bootstrap methods can
  be applied to heterogeneous $L_{2+\delta}$-\ned\ processes of size
  $-2(r-1)/(r-2)$ on a strong mixing sequence of size
  $-r(2+\delta)/(r-2)$, where $r > 2$ and $\delta >0$, when the
  original series has uniformly bounded $3r$-moments.  \citet{GoJ:03}
  relax these conditions to $L_{2+\delta}$-\ned\ of size $-1$ and
  $r+\delta$ moments for the original series, and size
  $-(2+\delta)(r+\delta)/(r-2)$ for the underlying mixing series.
  Both papers require that the expected block length grow with $n$ and
  be $o(n^{1/2})$.  \cite{GoP:11} discuss these issues
  further.}\footnote{An array $\{X_{nt}\}$ is an $L_{\rho}$-\ned\
  process on a mixing array $\{V_{nt}\}$ if
  \begin{equation}
    \| X_{nt} - \E(X_{nt}
    \mid V_{n,t-m},\dots,V_{n,t+m}) \|_{\rho} \leq d_{nt} v_m
  \end{equation} 
  with $v_m \to 0$ as $m \to \infty$ and $\{d_{nt}\}$ an array of
  positive constants.  It is of size $-\gamma$ if $v_m = O(m^{-\gamma
    - \delta})$ for all $\delta>0$.  Dropping the index ``$n$'' gives
  the series definition.  Note that strong and uniform mixing arrays are
  not required to be stationary.} But these conditions are still stronger than
required for a \clt\ to hold; \citet{Jon:97} has established the \clt\
under $L_2$-\ned\ with smaller size and moment
restrictions.\footnote{\citet{Jon:97} proves that the \clt\ holds for
  averages of $L_2$-\ned\ processes of size $-1/2$ on a strong mixing
  series of size $-r/(r-2)$, $r > 2$ and the original series having
  bounded $r$-moments.}  This paper shows that these block bootstrap
methods consistently estimate the distribution of the sample mean
under \citepos{Jon:97} assumptions, and show that an \fclt\ holds as
well.\footnote{\citet{Rad:96} proves consistency for the Moving Blocks
  Bootstrap for any stationary strong mixing sequence that satisfies
  the \clt. This paper uses a similar method of proof to his, but also
  accommodates nonstationary sequences and the Stationary Bootstrap.}
It also relaxes \citepos{GoW:02} and \citepos{GoJ:03} requirement that
the expected block length be $o(n^{1/2})$ to the original papers'
requirement that it be $o(n)$.

The proof exploits the conditional independence of the blocks in each
bootstrap.  Each bootstrap proceeds by drawing blocks of $M$
consecutive observations from the original time series, and then
pasting these blocks together to create the new bootstrap time series.
The Moving Blocks bootstrap does exactly that; the Circular Block
bootstrap ``wraps'' the observations, so that $(X_{n-1}, X_n, X_1,
X_2)$, for example, is a possible block of length four (letting $X_t$
denote the original time series).  The Stationary Bootstrap wraps the
observations and also draws $M$ at random for each block;
\citet{PoR:94} suggest drawing $M$ from the geometric distribution.
As the name suggests, the series produced by the Stationary Bootstrap
are strictly stationary, while those produced by the other methods are
not.  Although the Stationary Bootstrap was believed to be much less
efficient than other block bootstrap methods due to results of
\citet{Lah:99}, \citet{Nor:09} has shown that it is only slightly less
efficient than the other block-bootstrap methods discussed in this
paper, and has efficiency identical to that of the non-overlapping
block bootstrap.  Consequently, there has been renewed interest in the
Stationary Bootstrap since stationarity of the bootstrap samples can
be a useful property in theoretical research.  \cite{KrP:11} provides
a recent review of the bootstrap for time-series
processes,\footnote{Also see the discussion papers:
  \citet{Dah:11,GoP:11,Hor:11,JeM:11,KrP:11b}.} and \citet{GoP:11}
further discuss recent developments in block-bootstraps.

Theorem~\ref{thm:1} presents the main result, asymptotic normality of
the distribution of bootstrapped partial sums.  This paper adopts the
standard notation that $\E^{*}$, $\var^{*}$, etc. are the usual
operators with respect to the probability measure induced by the
bootstrap and will use explicit stochastic array notation 
for precision.  Also note that all results are presented for the
scalar case but generalize immediately to random vectors.  All of the
proofs are presented in the appendix; only proofs for the Stationary
Bootstrap are presented, since proofs for the other methods are
similar and easier to follow.  All limits are taken as $n \to \infty$
unless otherwise noted and $\lVert \cdot \rVert_r$ denotes the
$L_r$-norm.

\begin{thm}\label{thm:1}
  Suppose the following conditions hold.
  \begin{enumerate}
  \item $X_{nt}$ is $L_2$-\ned\ of size $-1/2$ on an array
    $V_{nt}$ that is either strong mixing of size $-r/(r-2)$ or
    uniform mixing of size $-r/2(r-1)$, with $r > 2$.  The
    \ned\ magnitude indices are denoted $\{d_{nt}\}$.
  \item The array $\mu_{nt} - \bar \mu_n$ is uniformly bounded where
    $\E X_{nt} = \mu_{nt}$ and $\bar{\mu}_n = n^{-1} \sum_{t=1}^n
    \mu_{nt}$.  Moreover, $\sqrt{n} \| \bar{X}_{n} - \bar\mu_n \|_2
    \to \sigma > 0$.
  \item There exists an array of positive real numbers $\{c_{nt}\}$
    such that $(X_{nt} - \mu_{nt})/c_{nt}$ is uniformly $L_r$-bounded
    and $c_{nt}$ and $d_{nt}/c_{nt}$ are uniformly bounded in $n$
    and~$t$.
  \item $X_{nt}^{*}$ is generated by the Stationary Bootstrap with
    geometric block lengths with success probability $p_n$, $p_n = c
    n^{-a}$ and $a,c \in (0,1)$, or by the Moving or Circular Block
    bootstrap with block length $M_n$ such that $M_n \sim n^a$ for 
    $a \in (0,1)$.  Let $M_{ni}$ be the block length of the $i$th
    block, $i=1,\dots,J_n$, and define $K_{n0} = 0$ and $K_{nj} =
    \sum_{i=1}^j M_{ni}$.
  \end{enumerate}
  Then
  \begin{equation}\label{eq:30}
    \sup_x \big\lvert \pr^{*}\big[\sqrt{n}(\bar X_{n}^{*} - \mu_{n}^{*}) \big/
    \sigma_n^{*} \leq x \big] - \Phi(x) \big\rvert \to^p 0,
  \end{equation}
  and
  \begin{equation}
    \label{eq:9}
    \sup_x \big\lvert \pr^{*}\big[
    \sqrt{n}(\bar X_{n}^{*} - \mu_{n}^{*}) \big/ \hat\sigma_n^{*}
    \leq x \big] - \Phi(x) \big\rvert \to^p 0,
  \end{equation}
  where $\Phi$ is the \textsc{cdf} of the Standard Normal
  distribution, $\sigma_n^{*2} = \var^{*}(\sqrt{n} \bar X_n^{*})$, and
    \begin{equation}
      \label{eq:12}
      \hat{\sigma}_n^{*2} = \tfrac{1}{n} \sum_{j=1}^{J_n}
      \Big\{\sum_{t=K_{n,j-1}+1}^{K_{nj}} (X_{nt}^{*} - \bar X_{n}^{*})\Big\}^2.
    \end{equation}
\end{thm}

Note that \citet{Jon:97} allows a little bit more flexibility in his
conditions on the array $\{c_{nt}\}$ \citep[see also][]{Dav:93};
essentially, he allows for a single set of blocks with the
maximal $\{c_{nt}\}$ over each block well-behaved, while this
paper requires this condition to hold for every possible partition of
blocks.  This additional restriction is required because the
Stationary Bootstrap will select the blocks randomly and is similar to
\citepos{JoD:00b} for the \fclt.

Consistency for the distribution of the sample mean follows as an
immediate corollary of Theorem~\ref{thm:1} and \citet[Theorem~2]{Jon:97}.

\begin{cor}\label{cor:1} 
  If the conditions of Theorem~\ref{thm:1} hold and $\hat{\sigma}^2_n$
  is a consistent estimator of $\sigma^2$ then
  \begin{equation}
    \label{eq:31}
    \sup_{x} \big\lvert \pr^{*}\big[
    \sqrt{n}(\bar X_{n}^{*} - \mu_{n}^{*}) / \sigma_{n}^{*}
    \leq x \big]
    - \pr\big[\sqrt{n}(\bar X_n - \bar{\mu}_n)/\hat{\sigma}_n \leq x \big] \big\rvert \to^p 0.
  \end{equation}
  and
  \begin{equation}
    \label{eq:11}
    \sup_{x} \big\lvert \pr^{*}\big[
    \sqrt{n} (\bar X_{n}^{*} - \mu_{n}^{*}) \big/ \hat\sigma^{*}_{n}
    \leq x \big]
    - \pr\big[\sqrt{n}(\bar X_n - \bar{\mu}_n)/\hat{\sigma}_n \leq x \big] \big\rvert \to^p 0.
  \end{equation}
\end{cor}

Corollary~\ref{cor:1} justifies using the Stationary Bootstrap to
conduct inference about $\bar \mu_n$ even though there may be
considerable heterogeneity.  Unlike \citet{GoW:02} and \citet{GoJ:03},
we are able to achieve consistency without
additional assumptions that ensure the heterogeneity is asymptotically
irrelevant, either because the deviations of $\mu_{nt}$ from $\bar
\mu_n$ are small or because there are a finite number of breaks that
occur in a neighborhood of the first observation.
The additional cost is the assumption that there must be an estimator
$\hat\sigma^2$ of $\sigma^2$ that is known to be consistent, and such an
estimator will typically require the heterogeneity in the mean to be
approximately known.

If the mean \textit{is} approximately constant, we can get a result
directly comparable to \citepos{GoW:02} and \citet{GoJ:03}, but with
the weaker size and moment conditions of Theorem~\ref{thm:1}.
Corollary~\ref{cor:2} formalizes this result and follows from
Corollary~\ref{cor:1} and Lemma~\ref{res:B}.
\begin{cor}\label{cor:2}
  Suppose that the conditions of Theorem~\ref{thm:1} hold and that
  $\sum_{t=1}^n (\mu_{nt} - \bar{\mu}_n)^2$ is $o(n^{1/2})$.  
  Then $\sigma_n^{*2} \to^p
  \sigma^2$, $\hat{\sigma}_n^{*2} \to^p \sigma^2$, and
\end{cor}
  \begin{equation}
    \sup_{x} \big\lvert \pr^{*}\big[
    \sqrt{n}(\bar X_{n}^{*} - \mu_{n}^{*})
    \leq x \big]
    - \pr\big[\sqrt{n}(\bar X_n - \bar{\mu}_n) \leq x \big] \big\rvert \to^p 0.
  \end{equation}

For some uses, a \clt\ is not enough and an \fclt\ is desired.
Theorem~\ref{thm:1} can be extended to give an \fclt\ (using
arguments from \citealp{JoD:00b}), but we will see that extending 
Corollaries~\ref{cor:1} and~\ref{cor:2} is more problematic.
Theorem~\ref{thm:2} proves a version of Theorem~\ref{thm:1} for the
partial sum; under the same assumptions as Theorem~\ref{thm:1}, the
partial sum of the bootstrapped process obeys an \fclt\ and can be
used to derive critical values for other test statistics.

\begin{thm}\label{thm:2}
  Suppose that the conditions of Theorem~\ref{thm:1} hold, let $W$ be
  standard Brownian Motion, and define
  \begin{equation}
    \label{eq:7}
    W_n^{*}(\gamma) = \tfrac{1}{\sqrt{n}}
    \sum_{t=1}^{\lfloor \gamma n \rfloor} (X^*_{nt} - \mu_n^{*}) / \sigma_{n}^{*}
  \end{equation}
  and 
  \begin{equation}
    \label{eq:7}
    \hat{W}_n^{*}(\gamma) = \tfrac{1}{\sqrt{n}}
    \sum_{t=1}^{\lfloor \gamma n \rfloor} (X^*_{nt} - \mu_n^{*}) / \hat\sigma_n^{*}.
  \end{equation}
  Then $\pr^{*}[\dist(W_n^{*}, W) > \delta] \to 0$ i.p. and
  $\pr^{*}[\dist(\hat{W}_n^{*}, W) > \delta] \to 0$ i.p. for any
  positive $\delta$ and distance $\dist$ that metricizes weak
  convergence.
\end{thm}

Just as with Theorem~\ref{thm:1}, we have natural corollaries using
Theorem~\ref{thm:2} to approximate the behavior of partial sums of the
original series.  But applying this corollary requires much stronger
assumptions.  In Corollary~\ref{cor:1}, the
heterogeneity in $X_{nt}$ affects inference through the estimator of
the asymptotic variance, $\hat{\sigma}_n^2$; but if the researcher
has a consistent estimator of the variance, inference on the grand
mean, $\bar{\mu}_{n}$, can proceed as normal.
Assuming negligible heterogeneity in the means was unnecessary unless
we want to avoid estimating $\sigma^2$ directly.

But for the partial sum, a consistent estimator of the variance is not
enough.  If the mean is not approximately constant then only the full
summation is centered at the grand mean---the partial sums are
centered at partial averages of the $\mu_{nt}$ terms.  If the research
question concerns the individual $\mu_{nt}$'s then the bootstrap
approximation can be used, but otherwise it can not.

A second difference is that, if the $\mu_{nt}$ terms are approximately
constant as in Corollary~\ref{cor:2}, the covariance process must
still be consistently estimated.  The bootstrapped processes are
(essentially) homoskedastic, so they are unable to match patterns of
heteroskedasticity in the original series.  Those patterns must be
removed by studentizing both the original and bootstrap partial sums.
Other methods, such
as the Local Block Bootstrap \citep{PaP:02,DPP:03}, may be able to
capture this additional heterogeneity, but we do not pursue that
possibility further.

>>>>

\documentclass[12pt]{article}
\usepackage{amsfonts}
\usepackage{amsmath,amsthm,amssymb,graphicx,setspace,url,booktabs,tabularx}
\usepackage[sort,round]{natbib}
\usepackage[margin=1.25in]{geometry}
\usepackage[small]{caption}

@@

\newtheorem{thm}{Theorem}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{claim}[thm]{Claim}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{res}[thm]{Result}
\newtheorem{lema}{Lemma}[section]
\newtheorem{alg}{Algorithm}
\newtheorem{asmp}{Assumption}[section]

@@

\DeclareMathOperator{\const}{const}
\DeclareMathOperator{\cov}{cov}
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\eqd}{\overset{d}{=}}
\DeclareMathOperator{\ind}{1}
\DeclareMathOperator{\pr}{Pr}
\DeclareMathOperator{\sgn}{sgn}
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\vech}{vech}

@@

\newcommand{\clt}{\textsc{clt}}
\newcommand{\fclt}{\textsc{fclt}}
\newcommand{\hac}{\textsc{hac}}
\newcommand{\lln}{\textsc{lln}}
\newcommand{\ned}{\textsc{ned}}

\frenchspacing
\onehalfspacing

@@

\author{Gray Calhoun\thanks{Economics Department, Iowa State
    University, Ames, IA 50011. Telephone: (515) 294-6271.  Email:
    \texttt{gcalhoun@iastate.edu}. Web: \texttt{http://www.econ.iastate.edu/$\sim$gcalhoun}.}}
\title{Block bootstrap consistency\\under weak assumptions}
\date{\today}
\maketitle

\begin{abstract}\noindent
  This paper weakens the size and moment conditions needed for typical
  block bootstrap methods (i.e. the moving blocks, circular blocks,
  and stationary bootstraps) to be valid for the sample mean of
  Near-Epoch-Dependent functions of mixing processes; they are
  consistent under the weakest conditions that ensure the original
  process obeys a Central Limit Theorem \citep[those
    of][\textit{Econometric Theory}]{Jon:97}.  In doing so, this paper
  extends de Jong's method of proof, a blocking argument, to hold with
  random and unequal block lengths.  This paper also proves that
  bootstrapped partial sums satisfy a Functional CLT under the same
  conditions.

@@

\newpage\noindent Block bootstraps, e.g. the moving blocks
\citep{Kun:89,LiS:92}, circular block \citep{PoR:92}, and stationary
bootstraps \citep{PoR:94}, have become popular in Economics, partly
because they do not require the researcher to make parametric
assumptions about the data generating process.  They are valid under
general weak dependence and moment conditions.  Some recent papers
(\citealp{GoW:02}; \citealp{GoJ:03}) relax the dependence and moment
conditions of the original papers to fit with those commonly used in
Econometrics based on Near-Epoch-Dependence
(\ned).\footnote{\citet{GoW:02} show that these bootstrap methods can
  be applied to heterogeneous $L_{2+\delta}$-\ned\ processes of size
  $-2(r-1)/(r-2)$ on a strong mixing sequence of size
  $-r(2+\delta)/(r-2)$, where $r > 2$ and $\delta >0$, when the
  original series has uniformly bounded $3r$-moments.  \citet{GoJ:03}
  relax these conditions to $L_{2+\delta}$-\ned\ of size $-1$ and
  $r+\delta$ moments for the original series, and size
  $-(2+\delta)(r+\delta)/(r-2)$ for the underlying mixing series.
  Both papers require that the expected block length grow with $n$ and
  be $o(n^{1/2})$.}\footnote{An array $\{X_{nt}\}$ is an
  $L_{\rho}$-\ned\ process on a mixing array $\{V_{nt}\}$ if
  \begin{equation}
    \| X_{nt} - \E(X_{nt}
    \mid V_{n,t-m},\dots,V_{n,t+m}) \|_{\rho} \leq d_{nt} v_m
  \end{equation} 
  with $v_m \to 0$ as $m \to \infty$ and $\{d_{nt}\}$ an array of
  positive constants.  It is of size $-\gamma$ if $v_m = O(m^{-\gamma
    - \delta})$ for all $\delta>0$.  Dropping the index ``$n$'' gives
  the series definition.} But these conditions are still stronger than
required for a \clt\ to hold; \citet{Jon:97} has established the \clt\
under $L_2$-\ned\ with smaller size and moment
restrictions.\footnote{\citet{Jon:97} proves that the \clt\ holds for
  averages of $L_2$-\ned\ processes of size $-1/2$ on a strong mixing
  series of size $-r/(r-2)$, $r > 2$ and the original series having
  bounded $r$-moments.}  In this paper, I'll show that these block
bootstrap methods consistently estimate the distribution of the sample
mean under \citepos{Jon:97} assumptions, and show that an \fclt\ holds
as well.  I also relax \citepos{GoW:02} and \citepos{GoJ:03}
requirement that the expected block length be $o(n^{1/2})$ to the
original papers' requirement that it be $o(n)$.

The proof will exploit the conditional independence of the blocks in
each bootstrap.  Each bootstrap proceeds by drawing blocks of $M$
consecutive observations from the original time series, and then
pasting these blocks together to create the new bootstrap time series.
The moving blocks bootstrap does exactly that; the circular block
bootstrap ``wraps'' the observations, so that $(X_{n-1}, X_n, X_1,
X_2)$, for example, is a possible block of length four (letting $X_t$
denote the original time series).  The stationary bootstrap wraps the
observations and also draws $M$ at random for each block;
\citet{PoR:94} suggest drawing $M$ from the geometric distribution.
As the name suggests, the series produced by the stationary bootstrap
are strictly stationary, while those produced by the other methods are
not.

Theorem~\ref{res:C} presents our main result, asymptotic normality of
the distribution of bootstrapped partial sums.  This paper adopt the
standard notation that $\E^{*}$, $\var^{*}$, etc. are the usual
operators with respect to the probability measure induced by the
bootstrap and will use explicit stochastic array notation in this
paper for precision.  Also note that all results are presented for the
scalar case but generalize immediately to vector-valued random
variables.  All of the proofs are presented in the appendix; I only
present proofs for the stationary bootstrap, since proofs for the
other methods are similar and easier.  All limits are taken as $n \to
\infty$ unless otherwise noted.

\begin{thm}\label{res:C}
  Suppose the following conditions hold.
  \begin{enumerate}
  \item $X_{nt}$ is $L_2$-\ned\ of size $-1/2$ on an array $V_{nt}$
    that is either strong mixing of size $-r/(r-2)$ or uniform mixing
    of size $-r/2(r-1)$, with $r > 2$.  The \ned\ magnitude indices
    are denoted $\{d_{nt}\}$.
  \item $\E X_{nt} = \mu_{nt}$, $\mu_{nt} - \bar \mu_n$ is uniformly
    bounded, and $\sqrt{n} \| \bar{X}_{n} - \bar\mu_n \|_2 \to 1$.
  \item There exists an array of positive real numbers $\{c_{nt}\}$
    such that $\max_t c_{nt}$ is uniformly finite, $(X_{nt} -
    \mu_{nt})/c_{nt}$ is uniformly $L_r$-bounded, and $d_{nt}/c_{nt}$
    is uniformly bounded.
  \item $X_{nt}^{*}$ is generated by the stationary bootstrap with
    geometric block lengths with success probability $p_n$, $p_n = c
    n^{-a}$ and $a,c \in (0,1)$, or by the moving or circular block
    bootstrap with block length $M_n$ such that $M_n \to \infty$ and
    $M_n/n \to 0$.
  \end{enumerate}
  Define $\mathcal{B}$ as Brownian Motion and
  $\mathbb{B}_n^{*}(\gamma) = n^{-1/2} \sum_{t=1}^{\lfloor n \gamma
    \rfloor} (X_{nt}^{*} - \mu_{n}^{*})$.  Then
  $\sqrt{n}(\bar{X}_n^{*} - \mu_n^{*}) \to^d N(0,1)$ and
  $\mathbb{B}_n^{*} \to^d \mathcal{B}$.
\end{thm}

Theorem~\ref{res:C} assumes that the series have been normalized by
dividing by the square root of the (population) second moment of
$\sqrt{n} (\bar{X}_n - \bar{\mu}_n)$.  Sometimes, though, it's useful
to treat the variance of $X_{nt}$ as a random variable (call it
$\eta^2_n$).  In such cases the dependence and moment conditions
listed in Theorem~\ref{res:C} apply to the original array and not the
normalized array $\eta^{-1}_n X_{nt}$ as is clear from inspecting the
proofs in this paper and \citet{Jon:97}.\footnote{See also
  \citepos{HaH:80} Theorem 3.2, Corollary 3.1, and following remarks.}
Deriving bounds for $\eta_n^{-1} X_{nt}$ from primitive conditions on
$X_{nt}$ would be difficult.

Note that \citet{Jon:97} allows a little bit more flexibility in his
conditions on the array $\{c_{nt}\}$ \citep[see also][]{Dav:93};
essentially, he allows there to be a single set of blocks with the
maximal $\{c_{nt}\}$ over each block well-behaved, while this
paper requires that his condition hold for every possible partition of
blocks.  This additional restriction is required because the
stationary bootstrap will select the blocks randomly.

Consistency for the sample mean follows as an immediate corollary of
Theorem~\ref{res:C} and \citet[Theorem~2]{Jon:97}.

\begin{cor}\label{res:E} 
  Define $\bar{\mu}_n = n^{-1} \sum_{t=1}^n \mu_{nt}$.  If the
  conditions of Theorem~\ref{res:C} hold then $\sqrt{n}(\bar{X}_n^{*}
  - \mu_n^{*})$ and $\sqrt{n}(\bar{X}_n - \bar{\mu}_n)$ converge in
  distribution to the same limit.
\end{cor}

Corollary~\ref{res:E} justifies using the stationary bootstrap to
conduct inference about $\bar \mu_n$ even though there is considerable
heterogeneity.  This conclusion is also present in \citepos{GoW:02}
and \citepos{GoJ:03} papers, but those authors introduce and emphasize
additional assumptions that ensure the heterogeneity is asymptotically
irrelevant, either because the deviations of $\mu_{nt}$ from $\bar
\mu_n$ are small or because there are a finite number of breaks that
occur in a neighborhood of the first observation.  Such assumptions
are unnecessary.

But heterogeneity rules out a version of Corollary~\ref{res:E} for the
partial sum.  It is not hard to see that under the
assumptions of Theorem~\ref{res:C},
\begin{equation}
  n^{-1/2} \sum_{t=1}^{\lfloor \gamma n \rfloor} \Big(X_{nt}
  - (\gamma n)^{-1} \sum_{s=1}^{\lfloor \gamma n \rfloor}
  \mu_{nt}\Big) \to^d \mathcal{B}_{\xi},
\end{equation}
where $\mathcal{B}_{\xi}$ is a Gaussian process and $\xi$ captures the
asymptotic variance of the partial sums.  It would be uncommon for a
researcher to be interested in these particular random variables; one
would typically be more interested in $\bar \mu_n$ than $(\gamma
n)^{-1} \sum_{s=1}^{\lfloor \gamma n \rfloor} \mu_{nt}$, but these
bootstrap methods do not allow us to approximate the distribution of
$(\gamma n)^{-1/2} \sum_{t=1}^{\lfloor \gamma n \rfloor} (X_{nt} -
\bar \mu_n)$.  The same discussion applies if $\mu_{nt} = \bar{\mu}_n$
for all $t$ but $\mathcal{B}_{\xi} \neq \mathcal{B}$.  Other methods,
such as the local block bootstrap \citep{PaP:02,DPP:03}, may be able to
capture this heterogeneity, but we do not pursue that possibility
further.

For completeness, I'll present a consistency result under the
additional condition that the heterogeneity vanishes asymptotically
\citep[similar to Assumption 2.2 of][]{GoW:02}.  This result is a
corollary of Theorem~\ref{res:C} and \citet[Theorem 3.1]{JoD:00b}.

\begin{cor}\label{res:F} Let $\mathbb{B}_n(\gamma) = n^{-1/2} \sum_{t=1}^{\lfloor n \gamma \rfloor} (X_{nt} -
  \bar{\mu}_{n})$ and suppose the conditions of Theorem~\ref{res:C}
  hold, $\sup_t |\mu_{nt} - \bar{\mu}_n| \to 0$, and
  \begin{equation}
    n^{-1} \sum_{s,t=1}^{\lfloor \gamma n \rfloor} \E (X_{ns} -
    \mu_{ns}) (X_{nt} - \mu_{nt}) \to \gamma
  \end{equation}
  for all $\gamma \in [0,1]$.  Then $\mathbb{B}_n$ and
  $\mathbb{B}_n^{*}$ converge in distribution to the same limit.
\end{cor}

We can develop some more intuition by looking at the argument behind
Theorem~\ref{res:C}.  For concreteness, let $X_{n1},\dots,X_{nn}$
denote the original array and let $X_{n1}^{*},\dots,X_{nn}^{*}$ a
hypothetical series generated by the stationary bootstrap.
Conditional on the number of blocks, $N_n$, and the block lengths,
$M_{n1},\dots,M_{nN_n}$, we can see that
\begin{equation}
  \sqrt{n} \bar{X}^{*}_n = \frac{1}{\sqrt{N_n}} \sum_{i=1}^{N_n}
  \Bigg\{\sqrt{\frac{N_n}{n}} \sum_{t=K_{n,i-1}+1}^{K_{n,i}} X_{nt}^{*} \Bigg\},\qquad\text{with}\quad K_{n0} = 0,
  \quad K_{ni} = \sum_{j=1}^i M_{nj},
\end{equation} 
is the sum of $N_n$ independent random variables.  So $\sqrt{n}
(\bar{X}^{*}_n - \bar{X}_n)$ should obey a \clt\ under very weak moment
conditions and satisfy a relationship like
\begin{equation}
  \label{eq:10}
  \frac{\bar{X}_n^{*} - \mu_n^{*}}{\var^{*}(\bar{X}_n^{*} \mid
    M_{n1},\dots,M_{n,N_n})} \to^{d^{*}} N(0,1).
\end{equation}

This argument applies directly to the moving blocks and circular
bootstraps, since $N_n$ and $M_{ni}$ are deterministic.  But they are
stochastic for the stationary bootstrap and the randomness of $N_n$,
in particular, complicates a direct argument; conditioning on $N_n$
would require us to work with the distributions of $M_{nj}$ given
$M_{n1} + \cdots + M_{n,N_n} = n$.  This information should not matter
in the limit, though, and so, as a first step, I'll show that we can
replace $N_n$ with its expected value without affecting the asymptotic
distribution induced by the bootstrap.

\begin{lem}\label{res:A}
  Consider the following procedure:
  \begin{enumerate}
  \item Draw $M_{n1},\dots,M_{n,\lfloor n p_n \rfloor}$ independently from
    the geometric distribution with success probability $p_n$, let
    $m_n = \sum_{i=1}^{\lfloor n p_n \rfloor} M_{ni}$, and let
    $\mathcal{M}_n = (M_{n1},\dots,M_{n,\lfloor n p_n \rfloor})$.
  \item Draw $\lfloor n p_n \rfloor$ blocks from $X_{n1},\dots,X_{nn}$
    (with the observations ``wrapped'' as in the stationary
    bootstrap), with $M_{ni}$ the length of the $i$th block.
  \end{enumerate}
  Let $\{X_{nt}^{**}\}$ be a hypothetical array generated by this
  procedure, let $\{X_{nt}^{*}\}$ be an array generated by the
  stationary bootstrap, define $\mu^{*}_n = \mu_n^{**} = \bar X_n$,
  and define $\mathbb{B}_n^{**}(\gamma) = m_n^{-1/2}
  \sum_{t=1}^{\lfloor \gamma m_n \rfloor} (X_{nt}^{**} - \mu_n^{**})$.
  Under the conditions of Theorem~\ref{res:C}, $\mathbb{B}_n^{*}$ and
  $\mathbb{B}_n^{**}$ converge to the same limit.
\end{lem}

It is easy to see that if $X_{n1}^*,\dots,X_{nn}^*$ is a hypothetical
sequence generated by the moving or circular blocks bootstrap with
block length $M_n$ and $X_{n1}^{**},\dots,X_{n,m_n}^{**}$ is a
bootstrap generated as in Lemma~\ref{res:A} but with $\lfloor n/M_n
\rfloor$ blocks of length $M_n$, then the conclusion of
Lemma~\ref{res:A} still holds.

Given Lemma~\ref{res:A} and the preceding discussion, consistency for
the distribution of the sample mean can be expected to hold if
\begin{equation}\label{eq:5}
  \frac{\bar{X}^{**}_n - \mu_n^{**}}{\var^{**}(\bar{X}_n^{**} \mid
    \mathcal{M}_n)} \to^{d^{**}} N(0,1)
\end{equation}
and $\var^{**}(m_{n}^{1/2} \bar X_n^{**} \mid \mathcal
\mathcal{M}_n) \to^p 1$.  Lemma~\ref{res:B} establishes a stronger
result necessary for partial sums.

\begin{lem}\label{res:B}  
  If the conditions of Theorem~\ref{res:C} hold then
  $\var^{**}(m_n^{-1/2} \sum_{t=1}^{\lfloor \gamma m_n \rfloor}
  X_{nt}^{**} \mid \mathcal{M}_n) \to^p \gamma$ for all
  $\gamma \in [0,1]$.
\end{lem}

The crux of Lemma~\ref{res:B} is the recognition that
$\var^{**}(m_n^{1/2} \bar{X}_n^{**} \mid \mathcal{M}_n)$ can be written as an
average over squared blocks of consecutive observations of the
original series;\footnote{Obviously, this argument needs to hold for
  the partial sums as well, but I'll just discuss the overall average
  for a clearer presentation.} i.e.
\begin{equation}
  \var^{**}(m_n^{1/2} \bar{X}^{**}_n \mid \mathcal{M}_n) = n^{-1}
  \sum_{\tau=1}^n \sum_{j=1}^{\lfloor n p_n \rfloor} Y_{n, \tau, j}^2
\end{equation}
where $Y_{n, \tau, j} = n^{-1/2}\sum_{t=I_{n,\tau,j-1}+1}^{I_{n,\tau,j}}
(X_{nt} - \bar{X}_n)$ and $0 = I_{n,\tau,0} < I_{n,\tau,1} < \cdots <
I_{n,\tau,\lfloor n p_n \rfloor} = m_n$ (the specifics are given in the
proof and use different notation).  Standard arguments from
\citet{Jon:97} establish that $\sum_j Y_{n,\tau,j}^2 \to^p 1$ for each
$\tau$.

As a final point, the consistency of the bootstrap variance 
follows as a corollary to Lemmas~\ref{res:A} and~\ref{res:B}.

\begin{cor}\label{res:D}
  If the conditions of Theorem~\ref{res:C} hold then
  $\var^{*}(n^{1/2} \bar{X}_n^{*}) = \var(n^{1/2} \bar{X}_n) + o_p$.
\end{cor}
